---
title: "The pgainsim users guide"
author: "Nora Scherer, Pascal Schlosser"
package: "pgainsim v1.1"
output: 
  rmarkdown::html_document:
    md_extensions: "-autolink_bare_uris"
vignette: >
  %\VignetteIndexEntry{The pgainsim users guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}  
---

```{r setup, cache = F, echo = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

# Introduction

The pgainsim package allows after performing genome-wide association studies
under an additive, recessive and dominant model for a quantitative trait
to determine study specific critical values when one of the three models
is more informative than the other ones for a quantitative trait locus.
First, study specific p-gain values are simulated. Second, based on the
simulated values quantiles of the empirical density of the p-gain are computed.
Finally, the quantiles are exported or interpolated and the critical values are obtained.

# Installation

To install this package from the github page of the Institute of Genetic Epidemiology - University of Freiburg, start R (version "4.0") and enter:

```{r install_package}
if (!requireNamespace("devtools", quietly = TRUE))
   install.packages("devtools")
if (!requireNamespace("rmarkdown", quietly = TRUE))
   install.packages("rmarkdown")
if (!requireNamespace("pgainsim", quietly = TRUE))
  devtools::install_github("genepi-freiburg/pgainsim",build_vignettes = TRUE)
```

# Example Code

The function `p_gain_simulation()` provides a dataset of simulated p-gains of different modes
using `pgain_types` (rec, dom, add), `AFs` (vector with different allele frequencies),
`n_study` (study size), `n` (number of random draws) as input while parallizing across `cores` (number of CPU cores):

```{r simulation}
require(pgainsim)
sim_dat <- p_gain_simulation(pgain_types=c("add","rec"), AFs=c(0.3,0.5,0.7,0.9), n=10000L, cores=1L)
```

Because the distribution of the additive p-gain is the same for the allele frequencies
AF and 1-AF in the following functions the simulations for these two allele frequencies
are combined for the additive p-gain.

Based on the simulated data, the empirical density of one p-gain type for different
allele frequencies can be plotted by the function `p_gain_density_plot()` where
`xlim` describes the range of the x-axis. Here we predefined the xlim range to zoom
in on the relevant part of the density, which in turn leads to a warning as a number
of data points is excluded from the plot.

```{r density_plot}
p_gain_density_plot(pgain_type="rec", sim_data=sim_dat, xlim=c(0,5), print_pdf=FALSE)
```

Based on the simulated data, the (1-0.05/#tests)-quantiles of the empirical density of the p-gain are
computed by the function `p_gain_quantiles()` where #tests is the number of parallel tests and
`n_tests` describes the maximal number of parallel tests:

```{r quantiles}
quantile <- p_gain_quantiles(n_tests=5L, sim_data=sim_dat)
```

For applications with extensive multiple testing the p-gain-quantiles can be interpolated
by a log-linear fit by use of the function `p_gain_quantile_fit()` where `test_number` describes
the number of parallel tests for which the p-gain threshold should be determined:

```{r quantiles_fit}
list_fits <- p_gain_quantile_fit(pgain_quantile=quantile, test_number=50L, print_pdf=FALSE)
```

By means of `start_vec` the starting estimates for the log-linear-fit for different p-gain types
and allele frequencies can be changed.

For details on the options of all functions please see `?pgainsim` and the corresponding
application note by Scherer et al. 2021 (under review).

# Session Info
```{r sessionInfo}
sessionInfo()
warnings()
```
